Semi-permeable vesicles: mathematical formulation, numerics and physics

### 2020-05-06 vesicle_code/docs/permeableVes/main.tex
fixed typo
worked on relaxation example

\newif\ifTikz
%\Tikztrue
\Tikzfalse

@@

\section{Problem formulation} \label{sc:formulate}
Consider a two-dimensional semi-permeable vesicle suspended in an
unbounded viscous fluid domain, subjected to an imposed flow
$\uu_\infty(\xx)$, for any $\xx \in \RR^2$.  Assume that the interior
and exterior fluids have the same viscosity $\mu$. Let $\xx$ be the
position of the interface $\gamma$, $\uu$ be the fluid velocity, and
$p$ be the pressure. In the vanishing Reynolds number limit, the
governing equations for the ambient fluid are
%
\begin{subequations}
\begin{alignat}{3}
  \label{eqn:governing}
  -\nabla p + \mu \triangle \uu &= 0, 
  &&\xx \in \RR^2 \setminus \gamma, \\
% 
  \nabla \cdot \uu &= 0,  &&\xx \in \RR^2 \setminus \gamma, \\
%
  \uu(\xx) &\rightarrow \uu_\infty(\xx),
    \qquad &&\norm{\xx} \rightarrow \infty.
\end{alignat}
\end{subequations}%
%
Letting $\nn$ be the outward normal to $\gamma$, $\jump{\cdot}$ denote
the jump across the interface, and $T$ denote the hydrodynamic stress
tensor, the boundary conditions along the vesicle are
\begin{align}
  \label{eqn:jump1}
  \jump{T\cdot\nn} &= \ff_\{mathrm{mem}, \\
  \label{eqn:jump2}
  \jump{\uu} &= 0,
\end{align}
where $\nn$ points outward to the vesicle, and $\ff_\mathrm{mem}$ is the
membrane force that includes a bending term that minimizes the Helfrich
energy, and a tension term that acts as a Lagrange multiplier to satisfy
the local inextensibility condition. The bending stiffness only sets the
time scale, so we let $k_b = 1$ throughout the paper. For ease of
implementation, we modify the bending tension and forces as
$\ff_\mathrm{mem} = \ff_\mathrm{ben} + \ff_\mathrm{ten}$, where
\begin{align}
  \label{eqn:membraneForces}
  \ff_\mathrm{ben} = -\xx_{ssss}, \qquad
  \ff_\mathrm{ten} = (\sigma \xx_s)_s.
\end{align}
However, the physical tension $\Lambda$ can be computed as a
post-processing step as
\begin{align}
  \Lambda = \sigma + 1.5 \kappa^2,
\end{align}
where $\kappa$ is the membrane curvature.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Area Dynamics}
Following the analysis of Veerapaneni et al.~\cite{vee-raj-bir-pur2009},
we compute the steady-state shape of a semi-permeable vesicle in a
quiescent flow. The steady-state shape is a stationary point of the
Lagrangian
\begin{align}
  \mathcal{L} = \frac{1}{2}\int_{\gamma} \kappa^2 \, ds +
    \Lambda \left(\int_{\gamma} ds  - L \right),
\end{align}
where $\Lambda$ is the constant tension that acts as a Lagrange
multiplier to satisfy the length conservation. Note that there is no
pressure Lagrange multiplier since the vesicle's area fluctuates. The
force acing on the membrane is found by taking the variation of
$\mathcal{L}$ with respect to $\gamma$, and by setting the total force
to zero, the steady state shape satisfies
\begin{align}
  \kappa_{ss} + \frac{1}{2}\kappa^3 - \Lambda \kappa = 0.
  \label{eqn:curvODE}
\end{align}
with initial conditions $\kappa(0) = \kappa_0$ and $\kappa_s(0) = 0$.
Integrating the differential equation~\eqref{eqn:curvODE} with respect
to arclength, we have
\begin{align}
  \frac{\kappa_s^2}{2} + \frac{\kappa^4}{8} - 
    \frac{\Lambda}{2}\kappa^2 = \kappa_0^4 - \frac{\Lambda}{2}\kappa_0^2.
  \label{eqn:curvODE2}
\end{align}
Assuming $\kappa_s \neq 0$, this differential equation is separable, and
can be solved analytically using Mathematica. For impermeable vesicles,
it is possible for the curvature to be
periodic~\cite{vee-raj-bir-pur2009}. However, the solution
of~\eqref{eqn:curvODE2} depends on the {\em EllipticF} function, and it
can not be made periodic. Therefore, it the only solution
of~\eqref{eqn:curvODE2} is $\kappa(s) = \kappa_0$, meaning that the
steady state shape is a circle. By the inextensibility condition, the
steady state vesicle has a radius of $L/2\pi$, so its curvature and
tension are
\begin{align}
  \kappa = \frac{2\pi}{L}, \: \Lambda = \frac{2\pi^2}{L^2}.
  \label{eqn:SSshape}
\end{align}
Note that in contrast to a impermeable vesicle, the final tension is
positive rather than negative.

In addition to computing the steady shape, we can compute the transient
area of the vesicle. Parameterizing the vesicle with $\xx$, its area is
\begin{align}
  A(t) = \frac{1}{2}\int_{\gamma} (\xx \cdot \nn)\, ds .
\end{align}
The derivative of $A(t)$ includes terms due to the time derivative of
$\xx \cdot \nn$, and a term due to interface
stretching~\cite{lai-tse-hua2008}. However, the stretching term is zero
because of the local inextensibility condition. Therefore,
\begin{align}
  \dot{A}(t) &= \frac{1}{2} \int_{\gamma} \pderiv{}{t} 
    (\xx \cdot \nn) \, ds = 
  \frac{1}{2} \int_{\gamma} (\dot{\xx} \cdot \nn)\, ds  + 
  \frac{1}{2} \int_{\gamma} (\xx \cdot \dot{\nn})\, ds.
\end{align}
The second integrand can be computed as
\begin{align}
  \int_\gamma (\xx \cdot \dot{\nn})\, ds 
    = \int_\gamma \left(\xx \cdot \dot{\xx}_s^\perp \right)
    = -\int_\gamma \left(\xx^\perp \cdot \dot{\xx}_s\right)\, ds 
    = \int_\gamma \left(\xx_s^\perp \cdot \dot{\xx}\right)\, ds 
    = \int_\gamma (\dot{\xx} \cdot \nn)\, ds.
\end{align}
Therefore,
\begin{align}
  \dot{A}(t) &= \int_\gamma (\dot{\xx} \cdot \nn)\, ds 
  = \int_\gamma \left(\beta (\ff \cdot \nn)\nn + \SS[\ff]\right) 
      \cdot \nn\, ds 
  = \beta \int_\gamma (\ff \cdot \nn)\, ds,
\end{align}
where the last equality follows from the incompressibility of the
single-layer potential and the divergence theorem.  Finally, using
integration by parts and the Frenet-Serret identities, we have
\begin{align}
  \dot{A}(t) &= \beta \int_\gamma \left( -\xx_{ssss} + 
    (\sigma \xx_s)_s \right) \cdot \nn \, ds \\
  &= -\beta \int_\gamma \left(-\xx_{sss} + \sigma \xx_s 
    \right) \cdot \kappa \xx_s \, ds \\
  &= -\beta \int_\gamma \left(\xx_{ss} \cdot 
    (\kappa \xx_s)_s + \kappa \sigma \right) \, ds \\
  &= -\beta \int_\gamma \left(\kappa (\xx_{ss} \cdot \xx_{ss}) + 
    \kappa_s (\xx_{ss} \cdot \xx_s) + \kappa \sigma \right) 
    \, ds \\
  &= -\beta \int_\gamma \left(\kappa^3 + \kappa \sigma \right) 
    \, ds.
\end{align}
Finally, using the relationship between the Lagrange multiplier
$\sigma$, and the tension $\Lambda$, the area of a semi-permeable
vesicle is governed by
\begin{align}
  \dot{A}(t) = \beta \int_\gamma \left(
    \frac{\kappa^3}{2} - \kappa \Lambda \right) \, ds.
  \label{eqn:areaROC}
\end{align}
Note that this is consistent with the steady state curvature and
tension~\eqref{eqn:SSshape} since these constant values result in the
integrand in equation~\eqref{eqn:areaROC} being zero.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Quiescent Flow}
We start by considering a semi-permeable vesicle in a quiescent flow.
For an impermeable vesicle, the vesicle relaxes to a steady-state shape
with negative and constant tension~\cite{kra-win-sei-lip1996}. However,
from equation~\eqref{eqn:SSshape}, we know that a semipermeable vesicle
will relax to a circle with positive constant tension. The area of a
semi-permeable vesicle is governed by~\eqref{eqn:areaROC}, so we expect
that the time to reach steady state scales with $\beta$. In
Figure~\ref{fig:ellipseArea}, we plot the area of a vesicle with
different permeability rates. The area is computed directly from the
vesicle shape (solid lines) and by integrating
equation~\eqref{eqn:areaROC} (marks) with a first-order method. We also
include black dashed lines indicating the times, $t_{\mathrm{SS}}$, that
the area of the vesicle first exceeds 12.4, meaning that its area is
within 1.3\% of the steady state area.  These times are reported in
Table~\ref{tbl:ellipseRelaxTime}, and we see that the relaxation time
does scale with $\beta$.

\begin{table}[htp]
  \centering
  \begin{tabular}{|C{1cm}|*{6}{C{2cm}}|}
    \hline
    $\beta$ & $10^{0}$ & $10^{-1}$ & $10^{-2}$ & 
             $10^{-3}$ & $10^{-4}$ & $10^{-5}$ \\
    $t_\mathrm{SS}$ & $6.45 \times 10^{0}$ & $2.00 \times 10^{1}$ & 
                      $9.61 \times 10^{1}$ & $8.36 \times 10^{2}$ & 
                      $8.15 \times 10^{3}$ & $8.03 \times 10^{4}$ \\
    ratio & --- & 3.10 & 4.81 & 8.71 & 9.74 & 9.85 \\
    \hline
  \end{tabular}
  \caption{\label{tbl:ellipseRelaxTime} The time, $t_{\mathrm{SS}}$, for
  a semi-permeable vesicle to reach steady state.}
\end{table}

\begin{figure}[htp]
\centering
%\ifTikz
\input{figures/ellipseArea.tikz}
%\fi
  \caption{\label{fig:ellipseArea} The area of an elliptical
  semi-permeable vesicle, with permeability rate $\beta$ in a quiescent
  flow. The solid lines compute the area directly from the vesicle
  shape, while the marks are computed by integrating
  equation~\eqref{eqn:areaROC}. The dashed black lines are the times
  when the vesicle's area is within 1\% of its maximum value of $4\pi$.}
\end{figure}

@@

\begin{figure}[htp]
  \centering
%  \ifTikz
  \input{figures/ellipseVelocity1.tikz}
  \input{figures/ellipseVelocity2.tikz}
  \input{figures/ellipseVelocity3.tikz}
%  \fi
  \caption{\label{fig:vesVelocity} The velocity contributions due to
  semi-permeability (red) and force balance (blue). The green dashed
  line is the velocity of an impermeable vesicle with an identical
  initial shape.}
\end{figure}

\begin{figure}[htp]
%  \ifTikz
  \input{figures/ellipseVelocityNorms.tikz}
%  \fi
  \caption{\label{fig:vesVelocityNorm} The max norm of the velocity due
  to the force balance (blue) and semi-permeability. Also included is
  the velocity of an impermeable vesicle (green). The three black lines
  correspond to the times reported in Figure~\ref{fig:vesVelocity}.}
\end{figure}

\todo[inline]{Finally, we tension and flux of a non-standard shape
vesicle.}

\begin{figure}[htp]
\begin{minipage}{0.40\textwidth}
\ifTikz
\input{figures/starTension.tikz}
\fi
\end{minipage}
\hfill
\begin{minipage}{0.55\textwidth}
  \centering
\includegraphics[width=0.32\textwidth]{figures/StarTensionTime1.pdf}
\includegraphics[width=0.32\textwidth]{figures/StarTensionTime2.pdf}
\includegraphics[width=0.32\textwidth]{figures/StarTensionTime3.pdf} \\
\includegraphics[width=0.32\textwidth]{figures/StarTensionTime4.pdf}
\includegraphics[width=0.32\textwidth]{figures/StarTensionTime5.pdf}
\includegraphics[width=0.32\textwidth]{figures/StarTensionTime6.pdf}
\end{minipage}
  \caption{\label{fig:starTension} The tension of the time steps shown
  in Figure~\ref{fig:starShape}. The tension at the time horizon $T=1$
  is also included. At this point, the vesicle is nearly circular and
  the tension is constant and positive. The constant is given by
  equation~\eqref{eqn:SSshape}, and this value corresponds to the
  dashed black line.}
\end{figure}

\begin{figure}[htp]
\begin{minipage}{0.40\textwidth}
\ifTikz
\input{figures/starFlux.tikz}
\fi
\end{minipage}
\hfill
\begin{minipage}{0.55\textwidth}
\includegraphics[width=0.32\textwidth]{figures/StarFluxTime1.pdf}
\includegraphics[width=0.32\textwidth]{figures/StarFluxTime2.pdf}
\includegraphics[width=0.32\textwidth]{figures/StarFluxTime3.pdf} \\
\includegraphics[width=0.32\textwidth]{figures/StarFluxTime4.pdf}
\includegraphics[width=0.32\textwidth]{figures/StarFluxTime5.pdf}
\includegraphics[width=0.32\textwidth]{figures/StarFluxTime6.pdf}
\end{minipage}
  \caption{\label{fig:starFlux} The flux of the time steps shown in
  Figure~\ref{fig:starShape}. The tension at the time horizon $T=1$ is
  also included. At this point, the vesicle is nearly circular and there
  is no flux.}
\end{figure}

%\begin{figure}[htp]
%  \input{figures/starFlux.tikz}
%  \caption{\label{fig:starFlux} The flux of the time steps shown in
%  Figure~\ref{fig:starShape}. The tension at the time horizon $T=1$ is
%  also included. At this point, the vesicle is nearly circular and their
%  is no flux.}
%\end{figure}

@@

\ifTikz
\input{figures/shearPhaseDiagram.tikz} 
\fi

@@

\ifTikz
\input{figures/extensionalPhaseDiagram.tikz}
\fi

@@

 \newpage
 \paragraph{\bf \Huge OLD STUFF}

%%%%%%%%%%%%%%
\section{{$\Theta-L$} formulation}
Here, we attempt to apply the formulation described by Sohn et
al.~\cite{soh-tse-li-voi-low2010} to a semi-permeable vesicle.  In the
semi-permeable formulation, the vesicle velocity is
\begin{align}
  \pderiv{\XX}{t} = \uu(\XX(t)) + \beta (\ff \cdot \nn) \nn.
\end{align}
In the {$\Theta-L$} variables, we require a normal velocity $V$, and a
tangential velocity $T$. Since the inextensibility condition is $T_s +
\kappa V = 0$, a condition for the vesicle velocity is
\begin{align}
  (\uu \cdot \ss)_s = -\kappa(\beta (\ff \cdot \nn) + 
      \uu \cdot \nn).
\end{align}
This equation can be solved following equations (32)--(39)
in~\cite{soh-tse-li-voi-low2010}, with the only difference being that
$\tilde{\uu}$ satisfies
\begin{align}
  (\tilde{\uu} \cdot \ss)_s + \kappa(\tilde{\uu} \cdot \nn) =
    -(\vv \cdot \ss)_s - \kappa(\vv \cdot \nn) 
    - \beta\kappa(\ff \cdot \nn).
\end{align}
Compared to equation (40), only the last term is new.

%%%%%%%%%%%%%%
Finally, we write the time stepping method in matrix form as
\begin{align}
  A \left(
    \begin{array}{c}
      \xx^{N+1} \\ \sigma^{N+1}
    \end{array}
  \right) = 
  \left(
    \begin{array}{c}
      \xx^{N} + \Delta t \uu_\infty(\xx^N) \\ \mathbf{1}
    \end{array}
  \right),
\end{align}
where
\begin{align}
 A = \left(
  \begin{array}{cc}
    I + \Delta t \beta P^N B^N + \Delta t \SS^N B^N & 
    -\Delta t P^N T^N - \Delta t \SS^N T^N \\
    D^N & 0
  \end{array}
  \right).
\end{align}

>>>>

\newif\ifTikz
\Tikztrue

@@

\section{Problem formulation} \label{sc:formulate}
Consider a two-dimensional semi-permeable vesicle suspended in an
unbounded viscous fluid domain, subjected to an imposed flow
$\uu_\infty(\xx)$, for any $\xx \in \RR^2$.  Assume that the interior
and exterior fluids have the same viscosity $\mu$. Let $\xx$ be the
position of the interface $\gamma$, $\uu$ be the fluid velocity, and
$p$ be the pressure. In the vanishing Reynolds number limit, the
governing equations for the ambient fluid are
%
\begin{subequations}
\begin{alignat}{3}
  \label{eqn:governing}
  -\nabla p + \mu \triangle \uu &= 0, 
  &&\xx \in \RR^2 \setminus \gamma, \\
% 
  \nabla \cdot \uu &= 0,  &&\xx \in \RR^2 \setminus \gamma, \\
%
  \uu(\xx) &\rightarrow \uu_\infty(\xx),
    \qquad &&\norm{\xx} \rightarrow \infty.
\end{alignat}
\end{subequations}%
%
Letting $\nn$ be the outward normal to $\gamma$, $\jump{\cdot}$ denote
the jump across the interface, and $T$ denote the hydrodynamic stress
tensor, the boundary conditions along the vesicle are
\begin{align}
  \label{eqn:jump1}
  \jump{T\cdot\nn} &= \ff_\mathrm{mem}, \\
  \label{eqn:jump2}
  \jump{\uu} &= 0,
\end{align}
where $\nn$ points outward to the vesicle, and $\ff_\mathrm{mem}$ is the
membrane force that includes a bending term that minimizes the Helfrich
energy, and a tension term that acts as a Lagrange multiplier to satisfy
the local inextensibility condition. The bending stiffness only sets the
time scale, so we let $k_b = 1$ throughout the paper. For ease of
implementation, we modify the bending tension and forces as
$\ff_\mathrm{mem} = \ff_\mathrm{ben} + \ff_\mathrm{ten}$, where
\begin{align}
  \label{eqn:membraneForces}
  \ff_\mathrm{ben} = -\xx_{ssss}, \qquad
  \ff_\mathrm{ten} = (\sigma \xx_s)_s.
\end{align}
However, the physical tension $\Lambda$ can be computed as a
post-processing step as
\begin{align}
  \Lambda = \sigma + 1.5 \kappa^2,
\end{align}
where $\kappa$ is the membrane curvature.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Area Dynamics}
Following the analysis of Veerapaneni et al.~\cite{vee-raj-bir-pur2009},
we compute the steady-state shape of a semi-permeable vesicle in a
quiescent flow. The steady-state shape is a stationary point of the
Lagrangian
\begin{align}
  \mathcal{L} = \frac{1}{2}\int_{\gamma} \kappa^2 \, ds +
    \Lambda \left(\int_{\gamma} ds  - L \right),
\end{align}
where $\Lambda$ is the constant tension that acts as a Lagrange
multiplier to satisfy the length conservation. Note that there is no
pressure Lagrange multiplier since the vesicle's area fluctuates. The
force acing on the membrane is found by taking the variation of
$\mathcal{L}$ with respect to $\gamma$, and by setting the total force
to zero, the steady state shape satisfies
\begin{align}
  \kappa_{ss} + \frac{1}{2}\kappa^3 - \Lambda \kappa = 0.
  \label{eqn:curvODE}
\end{align}
with initial conditions $\kappa(0) = \kappa_0$ and $\kappa_s(0) = 0$.
Integrating the differential equation~\eqref{eqn:curvODE} with respect
to arclength, we have
\begin{align}
  \frac{\kappa_s^2}{2} + \frac{\kappa^4}{8} - 
    \frac{\Lambda}{2}\kappa^2 = \kappa_0^4 - \frac{\Lambda}{2}\kappa_0^2.
  \label{eqn:curvODE2}
\end{align}
Assuming $\kappa_s \neq 0$, this differential equation is separable, and
can be solved analytically using Mathematica. For impermeable vesicles,
it is possible for the curvature to be
periodic~\cite{vee-raj-bir-pur2009}. However, the solution
of~\eqref{eqn:curvODE2} depends on the {\em EllipticF} function, and it
can not be made periodic. Therefore, the only solution
of~\eqref{eqn:curvODE2} is $\kappa(s) = \kappa_0$, meaning that the
steady state shape is a circle. By the inextensibility condition, the
steady state vesicle has a radius of $L/2\pi$, so its curvature and
tension are
\begin{align}
  \kappa = \frac{2\pi}{L}, \: \Lambda = \frac{2\pi^2}{L^2}.
  \label{eqn:SSshape}
\end{align}
Note that in contrast to a impermeable vesicle, the final tension is
positive rather than negative.

In addition to computing the steady shape, we can compute the transient
area of the vesicle. Parameterizing the vesicle with $\xx$, its area is
\begin{align}
  A(t) = \frac{1}{2}\int_{\gamma} (\xx \cdot \nn)\, ds .
\end{align}
The derivative of $A(t)$ includes terms due to the time derivative of
$\xx \cdot \nn$, and a term due to interface
stretching~\cite{lai-tse-hua2008}. However, the stretching term is zero
because of the local inextensibility condition. Therefore,
\begin{align}
  \dot{A}(t) &= \frac{1}{2} \int_{\gamma} \pderiv{}{t} 
    (\xx \cdot \nn) \, ds = 
  \frac{1}{2} \int_{\gamma} (\dot{\xx} \cdot \nn)\, ds  + 
  \frac{1}{2} \int_{\gamma} (\xx \cdot \dot{\nn})\, ds.
\end{align}
The second integrand can be computed as
\begin{align}
  \int_\gamma (\xx \cdot \dot{\nn})\, ds 
    = \int_\gamma \left(\xx \cdot \dot{\xx}_s^\perp \right)
    = -\int_\gamma \left(\xx^\perp \cdot \dot{\xx}_s\right)\, ds 
    = \int_\gamma \left(\xx_s^\perp \cdot \dot{\xx}\right)\, ds 
    = \int_\gamma (\dot{\xx} \cdot \nn)\, ds.
\end{align}
Therefore,
\begin{align}
  \dot{A}(t) &= \int_\gamma (\dot{\xx} \cdot \nn)\, ds 
  = \int_\gamma \left(\beta (\ff \cdot \nn)\nn + \SS[\ff]\right) 
      \cdot \nn\, ds 
  = \beta \int_\gamma (\ff \cdot \nn)\, ds,
\end{align}
where the last equality follows from the incompressibility of the
single-layer potential and the divergence theorem.  Using integration by
parts and the Frenet-Serret identities, we have
\begin{align}
  \dot{A}(t) &= \beta \int_\gamma \left( -\xx_{ssss} + 
    (\sigma \xx_s)_s \right) \cdot \nn \, ds \\
  &= -\beta \int_\gamma \left(-\xx_{sss} + \sigma \xx_s 
    \right) \cdot \kappa \xx_s \, ds \\
  &= -\beta \int_\gamma \left(\xx_{ss} \cdot 
    (\kappa \xx_s)_s + \kappa \sigma \right) \, ds \\
  &= -\beta \int_\gamma \left(\kappa (\xx_{ss} \cdot \xx_{ss}) + 
    \kappa_s (\xx_{ss} \cdot \xx_s) + \kappa \sigma \right) 
    \, ds \\
  &= -\beta \int_\gamma \left(\kappa^3 + \kappa \sigma \right) 
    \, ds.
\end{align}
Finally, using the relationship between the Lagrange multiplier
$\sigma$, and the tension $\Lambda$, the area of a semi-permeable
vesicle is governed by
\begin{align}
  \dot{A}(t) = \beta \int_\gamma \left(
    \frac{\kappa^3}{2} - \kappa \Lambda \right) \, ds.
  \label{eqn:areaROC}
\end{align}
Note that this is consistent with the steady state curvature and
tension~\eqref{eqn:SSshape} since these constant values result in the
integrand in equation~\eqref{eqn:areaROC} being zero.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Quiescent Flow} 
We start by considering a semi-permeable vesicle in a quiescent flow.
For an impermeable vesicle, the vesicle relaxes to a steady-state shape
with negative and constant tension~\cite{kra-win-sei-lip1996}. However,
from equation~\eqref{eqn:SSshape}, we know that a semipermeable vesicle
will relax to a circle with positive constant tension. Since the area of
a semi-permeable vesicle is governed by~\eqref{eqn:areaROC}, we expect
that the time to reach steady state, and the time required for a small
amount of inflation, both scale with $\beta^{-1}$. We simulate a vesicle
with initial reduced area $0.65$ with several values of $\beta$.  In
Table~\ref{tbl:ellipseRelaxTimes}, we report the time, $t_\mathrm{SS}$,
required for the vesicle's area to be within 1.3\% of its maximum area,
and the time, $t_\mathrm{SP}$, required for the vesicle's area to exceed
1.3\% of its initial area. We see that both these time scales do scale
with $\beta^{-1}$. In Figure~\ref{fig:ellipseArea}, we plot the area of
a vesicle with different permeability rates. The area is computed
directly from the vesicle shape (solid lines) and by integrating
equation~\eqref{eqn:areaROC} (marks) with a first-order method. We also
include black dashed lines indicating the times, $t_{\mathrm{SS}}$, that
the vesicle is near its steady state circular shape.

\begin{table}[htp]
  \centering
  \begin{tabular}{|C{2cm}C{3cm}C{3cm}C{3cm}C{3cm}|}
    \hline
    $\beta$ & $t_\mathrm{SS}$ & ratio & $t_\mathrm{SP}$ & ratio \\
    \hline
    $10^{0}$  & $6.45 \times 10^{0}$ & ---  
              & $2.17 \times 10^{-2}$ & --- \\ 
    $10^{-1}$ & $2.00 \times 10^{1}$ & 3.10 & 
                $1.25 \times 10^{-1}$ & 5.76 \\
    $10^{-2}$ & $9.61 \times 10^{1}$ & 4.81 & 
                $1.59 \times 10^{0}$ & 12.72 \\
    $10^{-3}$ & $8.36 \times 10^{2}$ & 8.71 & 
                $2.19 \times 10^{1}$ & 13.77 \\
    $10^{-4}$ & $8.15 \times 10^{3}$ & 9.74 & 
                $2.14 \times 10^{2}$ & 9.77 \\
    $10^{-5}$ & $8.03 \times 10^{4}$ & 9.85 & 
                $2.20 \times 10^{3}$ & 10.14 \\
    \hline
  \end{tabular}
  \caption{\label{tbl:ellipseRelaxTimes} The time for a semi-permeable
  vesicle to reach relaxation time ($t_\mathrm{SS}$) and to begin
  inflating ($t_\mathrm{SP}$).}
\end{table}

\begin{figure}[htp]
\centering
\ifTikz
\input{figures/ellipseArea.tikz}
\fi
  \caption{\label{fig:ellipseArea} The area of an elliptical
  semi-permeable vesicle, with permeability rate $\beta$ in a quiescent
  flow. The solid lines compute the area directly from the vesicle
  shape, while the marks are computed by integrating
  equation~\eqref{eqn:areaROC}. The dashed black lines are the times
  when the vesicle's area is within 1\% of its maximum value of $4\pi$.}
\end{figure}

@@

\begin{figure}[htp]
  \centering
  \ifTikz
  \input{figures/ellipseVelocity1.tikz}
  \input{figures/ellipseVelocity2.tikz}
  \input{figures/ellipseVelocity3.tikz}
  \fi
  \caption{\label{fig:vesVelocity} The velocity contributions due to
  semi-permeability (red) and force balance (blue). The green dashed
  line is the velocity of an impermeable vesicle with an identical
  initial shape.}
\end{figure}

\begin{figure}[htp]
  \ifTikz
  \input{figures/ellipseVelocityNorms.tikz}
  \fi
  \caption{\label{fig:vesVelocityNorm} The max norm of the velocity due
  to the force balance (blue) and semi-permeability. Also included is
  the velocity of an impermeable vesicle (green). The three black lines
  correspond to the times reported in Figure~\ref{fig:vesVelocity}.}
\end{figure}

Finally, we consider a semi-permeable vesicle with initial regions of
positive and negative curvature. The permeability rate is $\beta=1$, and
several time steps are illustrated in Figure~\ref{fig:starShape}. As
expected, the steady state shape is circular. The vesicle undergoes
three general behaviors: (i) for $t \in (0,1.6 \times 10^{-2})$, the
vesicle has a net negative flux resulting in a loss of area; (ii) for $t
\in (1.6 \times 10^{-2},1.0 \times 10^{0})$, the vesicle has a net
positive flux and the tension most of the membrane is undergoing
negative tension; (iii) for $t \in (1.0 \times 10^{0},2.0 \times
10^{0})$, the vesicle is near its steady state and transitions to a
positive constant tension.

We compute the tension $\Lambda$ and the flux $(-\xx_{ssss} + (\sigma
\xx_s)_s) \cdot \nn$ and plot the results in
Figure~\ref{fig:starTensionFlux}.  In the first time interval, the
vesicle deflates. Near the points of largest negative curvature, there
is a large negative tension and large inflow (positive flux).
Correspondingly, near the points with largest positive curvature, there
is a small tension and outflow (negative flux). In the second time
interval, the tension is negative and the flux is positive nearly
everywhere. The regions of large negative curvature correspond to
regions with the largest negative tension and the most inflow. Finally,
in the last time interval, the tension transitions to a positive
constant function, the flux transitions to zero, and the result is a
steady state circular vesicle with positive constant tension.

\begin{figure}[htp]
    \ifTikz
    \input{figures/starRelaxation1.tikz}
    \quad
    \input{figures/starRelaxation2.tikz}
    \quad
    \input{figures/starRelaxation3.tikz}
    \quad
    \input{figures/starRelaxation4.tikz}
    \quad
    \input{figures/starRelaxation5.tikz}
    \fi
  \caption{\label{fig:starShape} An initially star-shaped semi-permeable
  vesicle in a shear quiescent flow with $\beta=1$. Note that the time
  steps are not equispaced.}
\end{figure}

\begin{figure}[htp]
  \centering
  \includegraphics[width=0.19\textwidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime1.pdf}
  \includegraphics[width=0.19\textwidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime2.pdf}
  \includegraphics[width=0.19\textwidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime3.pdf}
  \includegraphics[width=0.19\textwidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarTensionTime4.pdf}
  \includegraphics[width=0.19\textwidth,trim =2cm 5cm 0cm 5cm,clip=true]{figures/StarTensionTime5.pdf}

  \includegraphics[width=0.19\textwidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime1.pdf}
  \includegraphics[width=0.19\textwidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime2.pdf}
  \includegraphics[width=0.19\textwidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime3.pdf}
  \includegraphics[width=0.19\textwidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime4.pdf}
  \includegraphics[width=0.19\textwidth,trim =2cm 5cm 0cm 5cm, clip=true]{figures/StarFluxTime5.pdf}
  \caption{\label{fig:starTensionFlux} The vesicle tension (top) and
  flux (bottom) of the semi-permeable vesicle in quiescent
  flow with $\beta=1$.}
\end{figure}

@@

%\ifTikz
%\input{figures/shearPhaseDiagram.tikz} 
%\fi

@@

%\ifTikz
%\input{figures/extensionalPhaseDiagram.tikz}
%\fi

@@

% \newpage
% \paragraph{\bf \Huge OLD STUFF}
%
%%%%%%%%%%%%%%%
%\section{{$\Theta-L$} formulation}
%Here, we attempt to apply the formulation described by Sohn et
%al.~\cite{soh-tse-li-voi-low2010} to a semi-permeable vesicle.  In the
%semi-permeable formulation, the vesicle velocity is
%\begin{align}
%  \pderiv{\XX}{t} = \uu(\XX(t)) + \beta (\ff \cdot \nn) \nn.
%\end{align}
%In the {$\Theta-L$} variables, we require a normal velocity $V$, and a
%tangential velocity $T$. Since the inextensibility condition is $T_s +
%\kappa V = 0$, a condition for the vesicle velocity is
%\begin{align}
%  (\uu \cdot \ss)_s = -\kappa(\beta (\ff \cdot \nn) + 
%      \uu \cdot \nn).
%\end{align}
%This equation can be solved following equations (32)--(39)
%in~\cite{soh-tse-li-voi-low2010}, with the only difference being that
%$\tilde{\uu}$ satisfies
%\begin{align}
%  (\tilde{\uu} \cdot \ss)_s + \kappa(\tilde{\uu} \cdot \nn) =
%    -(\vv \cdot \ss)_s - \kappa(\vv \cdot \nn) 
%    - \beta\kappa(\ff \cdot \nn).
%\end{align}
%Compared to equation (40), only the last term is new.
%
%
%%%%%%%%%%%%%%%
%Finally, we write the time stepping method in matrix form as
%\begin{align}
%  A \left(
%    \begin{array}{c}
%      \xx^{N+1} \\ \sigma^{N+1}
%    \end{array}
%  \right) = 
%  \left(
%    \begin{array}{c}
%      \xx^{N} + \Delta t \uu_\infty(\xx^N) \\ \mathbf{1}
%    \end{array}
%  \right),
%\end{align}
%where
%\begin{align}
% A = \left(
%  \begin{array}{cc}
%    I + \Delta t \beta P^N B^N + \Delta t \SS^N B^N & 
%    -\Delta t P^N T^N - \Delta t \SS^N T^N \\
%    D^N & 0
%  \end{array}
%  \right).
%\end{align}
%
%\section{Area of a Vesicle}
%The area inside a single vesicle $\gamma$ parameterized with $\xx$ is
%\begin{align}
%  A(t) = \int_{\gamma} (\xx \cdot \nn)\, ds = 
%    \int_{0}^{2\pi} \left(\xx \cdot \pderiv{\xx}{\theta}^\perp\right)
%    \, d\theta.
%\end{align}
%Taking the time derivative of $A(t)$,
%\begin{align}
%  \dot{A}(t) &= \int_{0}^{2\pi} \left(
%    \dot{\xx} \cdot \pderiv{\xx}{\theta}^{\perp} + 
%    \xx \cdot \pderiv{\dot{\xx}}{\theta}^{\perp}\right) \\
%    \, d\theta.
%  &= \int_{0}^{2\pi} \left(
%    (\dot{\xx} \cdot \nn)\|\xx'(\theta)\| \right) d\theta +
%     \int_{0}^{2\pi}
%    \xx \cdot \frac{d}{dt}\left(\nn |\frac{d\xx}{d\theta}| \right)
%\end{align}
%
%
%
%\begin{align}
%  A(t) = \frac{1}{2}\int_{\gamma} (\xx \cdot \nn)\, ds = 
%    \frac{1}{2}\int_{0}^{2\pi} \left(\xx \cdot \pderiv{\xx}{\theta}^\perp\right)
%    \, d\theta.
%\end{align}
%Taking the time derivative of $A(t)$,
